### 2017-03-23 thesis/discussion/discussion.tex
Finished editing the draft thesis - clean things up

>>>>

\subsection{Use of \gls{svd} and transformation matrix as a valid method}
\label{sub:use_of_svd_and_transformation_matrix_as_a_valid_method}

From all of these results so far, one may question how robust or appropriate the use of \gls{svd} and transformation matrices are to generate metagenes in other data sets.
It appeared as though the CR data-derived obesity associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became insignificant.
One possible reason for this was that the obesity metagenes were not actually associated to obesity phenotype in the other data sets.
Another reason could be that the methodology used were not appropriate for investigating the association between the phenotype and its associated phenotype.

The appropriateness of the use of \gls{svd} and transformation matrices were examined with different genetic signatures: \gls{er} and \gls{pr} pathway associated genetic signatures from the \citet{Gatza2010a} study (covered in \cref{sec:results_for_er_and_pr_metagene_analysis}).
Three data sets (CR, FM and \gls{nzbc}) were used, as these data sets have both \gls{er} and \gls{pr} status information available for the patients.
\gls{svd} was first applied to the \gls{rma} normalised CR data to obtain the \gls{er} and \gls{pr} metagenes, along with the transformation matrices for each of the genetic signatures.
Both the \gls{er} and \gls{pr} metagenes showed good accordance with the gene expressions of the genetic signature in the heatmap, and also showed significant association with the patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).

Transformation matrices were applied to the FM and \gls{nzbc} data sets to generate the appropriate pathway metagenes in the data sets.
Both the \gls{er} and \gls{pr} pathway metagenes showed similar overall gene expression patterning in the heatmap as the CR data (\cref{sec:results_for_er_and_pr_metagene_analysis}).
Furthermore, both the \gls{er} and \gls{pr} pathway metagenes from these data sets showed significant association with the patient \gls{er} and \gls{pr} statuses, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
These results clearly showed that the general approach of using \gls{svd} and transformation matrix was appropriate for examining the association of the genetic signatures with the patient phenotype.

### 2017-03-23 thesis/misc/appendix.tex
Finished editing the draft thesis - clean things up

>>>>

\section{Results for \gls{er} and \gls{pr} metagene analysis}
\label{sec:results_for_er_and_pr_metagene_analysis}

To determine whether the results from all of the obesity metagene analyses were not showing any association with patient \gls{bmi} because of the \gls{svd}/transformation matrix approach used, the same approach was used with \gls{er} and \gls{pr} pathway associated genetic signatures from the \citet{Gatza2010a} study.
Firstly, both the \gls{er} and \gls{pr} metagenes were generated in the \gls{rma} normalised CR data set.
The generated metagenes were compared with the gene expressions of the corresponding pathway associated genetic signatures, and the association with the patient \gls{er} and \gls{pr} statuses were also examined.
As shown in \cref{fig:er_cr,fig:pr_cr}, both the \gls{er} and \gls{pr} metagenes in the CR data showed assocation with the gene expression pattern of the pathway, and also showed significant association with the patient \gls{er} and \gls{pr} statuses.

	\begin{figure}[htp!]
		\centering
		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_cr}\\
		\hfill
		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_cr}
		\caption[\acrshort{er} metagene in the CR data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the CR data.
		Scales for the heatmap is as described in previou figures.
		P-values  are calculated relative to the \gls{er}$^-$ group.}
		\label{fig:er_cr}
	\end{figure}

	\begin{figure}[htp!]
		\centering
		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cr}\\
		\hfill
		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cr}
		\caption[\acrshort{pr} metagene in the CR data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the CR cancer data.
		Scales and  p-values are as described in previous figures.}
		\label{fig:pr_cr}
	\end{figure}

The transformation matrices for the \gls{er} and \gls{pr} metagenes were created in the CR data, and was applied to the \gls{nzbc} and FM data sets (\gls{rma} normalised).
As clearly shown by \cref{fig:er_cris,fig:pr_cris,fig:er_fm,fig:pr_fm}, both the \gls{er} and \gls{pr} metagenes associated with the gene expressions of the corresponding pathway signature.
More importantly, both metagenes were significantly associated with the patient \gls{er} and \gls{pr} statuses, even though these metagenes were generated with the transformation matrices from the CR data.
These results show that the approach taken in this project does not affect the link between the genetic signature and the patient phenotype, if such link exists between them.

	\begin{figure}[htp!]
		\centering
		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_cris}\\
		\hfill
		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_cris}
		\caption[\acrshort{er} metagene in the \gls{nzbc} data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the \gls{nzbc} data.
		Scales for the heatmap is as described in previou figures.
		P-values  are calculated relative to the \gls{er}$^-$ group.}
		\label{fig:er_cris}
	\end{figure}

	\begin{figure}[htp!]
		\centering
		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cris}\\
		\hfill
		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cris}
		\caption[\acrshort{pr} metagene in the \gls{nzbc} data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the \gls{nzbc} cancer data.
		Scales and  p-values are as described in previous figures.}
		\label{fig:pr_cris}
	\end{figure}

	\begin{figure}[htp!]
		\centering
		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_fm}\\
		\hfill
		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_fm}
		\caption[\acrshort{er} metagene in the FM data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the FM data.
		Scales for the heatmap is as described in previou figures.
		P-values  are calculated relative to the \gls{er}$^-$ group.}
		\label{fig:er_fm}
	\end{figure}

	\begin{figure}[htp!]
		\centering
		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_fm}\\
		\hfill
		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_fm}
		\caption[\acrshort{pr} metagene in the FM data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the FM cancer data.
		Scales and  p-values are as described in previous figures.}
		\label{fig:pr_fm}
	\end{figure}
